Towards more realistic model representation of microbial population dynamics in complex environment

### 2020-03-31 SI_code_revised.Rmd
Tables numbering corrected

#Bacteria
bac<-as.data.frame(t(read.csv(file=c("bacteria.csv"), skip = 4)[,-1]))
colnames(bac)<-t(read.csv(file=c("bacteria.csv"), skip = 4)[,1])
##remove OTUs with all zeros
bac2<-bac[, colSums(bac!=0)>0]

bac_env<-as.data.frame(t(read.csv(file=c("bacteria.csv"), nrows = 4, header = F)[,-1]))
colnames(bac_env)<-t(read.csv(file=c("bacteria.csv"), nrows = 4, header = F)[,1])

bac_env$Structure<-recode(bac_env$Structure, 'Broth'="BROTH", 'Glass wool'="WOOL",
                          'Mixed glass'="GLASS")

#Fungi
fungi<-as.data.frame(t(read.csv(file=c("fungi.csv"), skip = 4)[,-1]))
colnames(fungi)<-t(read.csv(file=c("fungi.csv"), skip = 4)[,1])

##remove OTUs with all zeros
fungi2<-fungi[, colSums(fungi!=0)>0]

fungi_env<-as.data.frame(t(read.csv(file=c("fungi.csv"), nrows = 4, header = F)[,-1]))
colnames(fungi_env)<-t(read.csv(file=c("fungi.csv"), nrows = 4, header = F)[,1])

fungi_env$Structure<-recode(fungi_env$Structure, 'Broth'="BROTH", 'Glass wool'="WOOL",
                          'Mixed glass'="GLASS")

@@

```
##Statistical analysis
Linear Mixed-Effect models are used to evaluate the effect of substrate and the type of physical barrier on measured respiration rate, DNA and protein concentration. Because all three variables display significant variability over time, the time of measurement is treated as a random effect factor. Type of the barrier is nested within the substrate.

@@

```
$~$
$~$
$~$
##Microbial-explicit models analysis
###Both models are first calibrated against respiration rate
Calibration of both models is done in four steps. First, the model is calibrated against data across all treatments. Second, the model is calibrated for each level of incubation system complexity separately. Third, the model is calibrated for Glucose and Cellobiose treatments separately. Fourth, the model is calibrated for each experimental treatment separately. 
Goodness of correspondence between measurements and model simulations are calculated across all treatments. By doing so, the effect of experimental treatment is evaluated. If the model simulation fits the data significantly better when model parameters are estimated for different treatments separately, effect of this treatment is significant.

@@

####Monod model
```{r monod respiration calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/monod_r.R")
source("../Rscripts/monod_r_fix.R")
#across all treatments
monod_r1<-monod_r(data=d, FACT = 4)
#########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
##########################################################
#for different structures
monod_r2<-monod_r(data=d, FACT = 2)
monod_r2_fix<-monod_r_fix(data=d, FACT = 2)
#for different substrates
monod_r3<-monod_r(data=d, FACT = 1)
monod_r3_fix<-monod_r_fix(data=d, FACT = 1)
#for each separately  
monod_r4<-monod_r(data=d, FACT = 3)
monod_r4_fix<-monod_r_fix(data=d, FACT = 3)
###################################################
stopImplicitCluster()
###################################################
# monod_r1$goodness
# monod_r2$goodness
# monod_r3$goodness
# monod_r4$goodness

@@

kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('v', 'k', '$\\Y_{CUE}$', 'm'), row.names=T, caption = "_**Table S3:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_r_results_good[c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S4:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

@@

####Sub-microbial model
```{r Sub-microbial model respiration calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/two_pool_r.R")
source("../Rscripts/two_pool_r_fix.R")
#across all treatments
two_pool_r1<-two_pool_r(data=d, FACT = 4)
#################################################
no_cors<-detectCores()
cl<-makeCluster(no_cors)
registerDoParallel(cl)
#################################################
#for different structures
two_pool_r2<-two_pool_r(data=d, FACT = 2)
two_pool_r2_fix<-two_pool_r_fix(data=d, FACT = 2)
#for different substrates
two_pool_r3<-two_pool_r(data=d, FACT = 1)
two_pool_r3_fix<-two_pool_r_fix(data=d, FACT = 1)
#for each separately 
two_pool_r4<-two_pool_r(data=d, FACT = 3)
two_pool_r4_fix<-two_pool_r_fix(data=d, FACT = 3)
#################################################
stopImplicitCluster()
#################################################
# mean(c(two_pool_r4[[1]]$pars[["R_0"]], two_pool_r4[[2]]$pars[["R_0"]],
#      two_pool_r4[[3]]$pars[["R_0"]], two_pool_r4[[4]]$pars[["R_0"]],
#      two_pool_r4[[5]]$pars[["R_0"]], two_pool_r4[[6]]$pars[["R_0"]]))
# mean(c(two_pool_r4[[1]]$pars[["S_0"]], two_pool_r4[[2]]$pars[["S_0"]],
#      two_pool_r4[[3]]$pars[["S_0"]], two_pool_r4[[4]]$pars[["S_0"]],
#      two_pool_r4[[5]]$pars[["S_0"]], two_pool_r4[[6]]$pars[["S_0"]]))

@@

#results of the previous iteration (not run)
# two_pool_r_results_par<-rbind(two_pool_r1[[1]]$pars[c(1:5)],
#                    two_pool_r2_fix[[1]]$pars, two_pool_r2_fix[[2]]$pars,
#                    two_pool_r2_fix[[3]]$pars,
#                    two_pool_r3_fix[[1]]$pars, two_pool_r3_fix[[2]]$pars,
#                    two_pool_r4_fix[[1]]$pars, two_pool_r4_fix[[2]]$pars,
#                    two_pool_r4_fix[[3]]$pars, two_pool_r4_fix[[4]]$pars,
#                    two_pool_r4_fix[[5]]$pars, two_pool_r4_fix[[6]]$pars)
# two_pool_r_results_good<-rbind(two_pool_r1$goodness,
#                               two_pool_r2$goodness,
#                               two_pool_r3$goodness,
#                               two_pool_r4$goodness)
# write.csv(two_pool_r_results_par, file = c("../two_pool_r_results_par.csv"), row.names = F)
# write.csv(two_pool_r_results_good, file = c("../two_pool_r_results_good.csv"), row.names = F)

@@

kable(t.s3.1, digits = c(2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S7:** Sub-microbial model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(two_pool_r_results_good[, c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S8:** Goodness of fit of Sub-microbial model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

@@

###Both models are calibrated against respiration rate, proteins and DNA concentration in the cells.
Both models are calibrated against all data we measured. In addition to model parameters reported in tables above, conversion factors between proteins/DNA concentration and microbial biomass are estimated. For Monod model, initial concentration of microbial biomass is defined as an initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and microbial biomass. For Sub-microbial model, initial concentration of Structures is defined similarly - i.e. initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and Structures. However, initial concetration of Reserves cannot be defined because initial concentration of proteins is not know. Therefore, one more parameter is defined. This parameter is denoted as RS_0 in the respective function and define initial concentration of Reserves per unit of Structures. This parameter has to be fixed across treatments to reflect the reality.
In addition, Monod model is calibrated against respiration rate and proteins or DNA concentration separately. When calibrated againnst protein concentration, initial concentration of microbial biomass is estimated and fixed across treatments.

####Monod model
```{r Monod model - biomass components, eval=F, echo=T}
#Proteins
#Loading the function 
source("../Rscripts/monod_prot.R")
source("../Rscripts/monod_prot_fix.R")
#across all treatments
monod_prot1<-monod_prot(data=d, FACT = 4)
###############################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###############################################################
#for different structures
monod_prot2<-monod_prot(data=d, FACT = 2)
monod_prot2_fix<-monod_prot_fix(data=d, FACT = 2)
#for different substrates
monod_prot3<-monod_prot(data=d, FACT = 1)
monod_prot3_fix<-monod_prot_fix(data=d, FACT = 1)
#for each separately 
monod_prot4<-monod_prot(data=d, FACT = 3)
monod_prot4_fix<-monod_prot_fix(data=d, FACT = 3)
###############################################################
stopImplicitCluster()
###############################################################
monod_prot1$goodness
monod_prot2_fix$goodness
monod_prot3_fix$goodness
monod_prot4_fix$goodness

@@

kable(t.s4.1, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S9:** Monod model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_all_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S10:** Goodness of fit of Monod model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 3) %>%
  group_rows("Structure", 4, 6) %>%
  group_rows("Substrate", 7, 9) %>%
  group_rows("Each", 10, 12)

@@

kable(t.s4.2, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S11:** Monod model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_prot_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S12:** Goodness of fit of Monod model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8)

@@

kable(t.s4.3, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S13:** Monod model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_DNA_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S14:** Goodness of fit of Monod model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8)

@@

####Sub-microbial model
Not all model parameters vary significantly across treatments. Parameters that doesn't show any variability across treatments are fixed.
```{r Sub-microbial model calibration, eval=F, echo=T}
#Loading the function 
source("./Rscripts/two_pool_allx.R")
source("../Rscripts/two_pool_allx_fix.R")
#across all treatments
two_pool_allx_out1<-two_pool_allx(data=d, FACT = 4)

@@

# write.csv(two_pool_all_results_par, file = c("../two_pool_all_results_par.csv"), row.names = F)
# write.csv(two_pool_all_results_good, file = c("../two_pool_all_results_good.csv"), row.names = F)

@@

kable(t.s6.1, digits = c(0, 2, 2, 2, 3, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ","v", 'k', 'f', 'm', 'Y', '$\\B_{R-Proteins}$', '$\\B_{S-Proteins}$', '$\\B_{S-DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S21:** Sub-microbial model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(two_pool_all_results_good[, c(1, 4:7)], digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S21:** Goodness of fit of Sub-microbial model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 3) %>%
  group_rows("Structure", 4, 6) %>%
  group_rows("Substrate", 7, 9) %>%
  group_rows("Each", 10, 12)

>>>>

@@

## Statistical analysis

Linear Mixed-Effect models are used to evaluate the effect of substrate and the type of physical barrier on measured respiration rate, DNA and protein concentration. Because all three variables display significant variability over time, the time of measurement is treated as a random effect factor. Type of the barrier is nested within the substrate.

@@

```
$~$
$~$
$~$

## Microbial-explicit models analysis

### Both models are first calibrated against respiration rate

Calibration of both models is done in four steps. First, the model is calibrated against data across all treatments. Second, the model is calibrated for each level of incubation system complexity separately. Third, the model is calibrated for Glucose and Cellobiose treatments separately. Fourth, the model is calibrated for each experimental treatment separately. 
Goodness of correspondence between measurements and model simulations are calculated across all treatments. By doing so, the effect of experimental treatment is evaluated. If the model simulation fits the data significantly better when model parameters are estimated for different treatments separately, effect of this treatment is significant.

@@

#### Monod model

```{r monod respiration calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/monod_r.R")
source("../Rscripts/monod_r_fix.R")
#across all treatments
monod_r1<-monod_r(data=d, FACT = 4)
#########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
##########################################################
#for different structures
monod_r2<-monod_r(data=d, FACT = 2)
monod_r2_fix<-monod_r_fix(data=d, FACT = 2)
#for different substrates
monod_r3<-monod_r(data=d, FACT = 1)
monod_r3_fix<-monod_r_fix(data=d, FACT = 1)
#for each separately  
monod_r4<-monod_r(data=d, FACT = 3)
monod_r4_fix<-monod_r_fix(data=d, FACT = 3)
###################################################
stopImplicitCluster()
###################################################
# monod_r1$goodness
# monod_r2$goodness
# monod_r3$goodness
# monod_r4$goodness

@@

kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('v', 'k', '$\\Y_{CUE}$', 'm'), row.names=T, caption = "_**Table S2:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_r_results_good[c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S3:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

@@

#### Sub-microbial model

```{r Sub-microbial model respiration calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/two_pool_r.R")
source("../Rscripts/two_pool_r_fix.R")
#across all treatments
two_pool_r1<-two_pool_r(data=d, FACT = 4)
#################################################
no_cors<-detectCores()
cl<-makeCluster(no_cors)
registerDoParallel(cl)
#################################################
#for different structures
two_pool_r2<-two_pool_r(data=d, FACT = 2)
two_pool_r2_fix<-two_pool_r_fix(data=d, FACT = 2)
#for different substrates
two_pool_r3<-two_pool_r(data=d, FACT = 1)
two_pool_r3_fix<-two_pool_r_fix(data=d, FACT = 1)
#for each separately 
two_pool_r4<-two_pool_r(data=d, FACT = 3)
two_pool_r4_fix<-two_pool_r_fix(data=d, FACT = 3)
#################################################
stopImplicitCluster()
#################################################
# mean(c(two_pool_r4[[1]]$pars[["R_0"]], two_pool_r4[[2]]$pars[["R_0"]],
#      two_pool_r4[[3]]$pars[["R_0"]], two_pool_r4[[4]]$pars[["R_0"]],
#      two_pool_r4[[5]]$pars[["R_0"]], two_pool_r4[[6]]$pars[["R_0"]]))
# mean(c(two_pool_r4[[1]]$pars[["S_0"]], two_pool_r4[[2]]$pars[["S_0"]],
#      two_pool_r4[[3]]$pars[["S_0"]], two_pool_r4[[4]]$pars[["S_0"]],
#      two_pool_r4[[5]]$pars[["S_0"]], two_pool_r4[[6]]$pars[["S_0"]]))

@@

#results of the previous iteration (not run)
# two_pool_r_results_par<-rbind(two_pool_r1[[1]]$pars[c(1:5)],
#                    two_pool_r2_fix[[1]]$pars, two_pool_r2_fix[[2]]$pars,
#                    two_pool_r2_fix[[3]]$pars,
#                    two_pool_r3_fix[[1]]$pars, two_pool_r3_fix[[2]]$pars,
#                    two_pool_r4_fix[[1]]$pars, two_pool_r4_fix[[2]]$pars,
#                    two_pool_r4_fix[[3]]$pars, two_pool_r4_fix[[4]]$pars,
#                    two_pool_r4_fix[[5]]$pars, two_pool_r4_fix[[6]]$pars)
# two_pool_r_results_good<-rbind(two_pool_r1$goodness,
#                               two_pool_r2$goodness,
#                               two_pool_r3$goodness,
#                               two_pool_r4$goodness)
# write.csv(two_pool_r_results_par, file = c("./two_pool_r_results_par.csv"), row.names = F)
# write.csv(two_pool_r_results_good, file = c("./two_pool_r_results_good.csv"), row.names = F)

@@

kable(t.s3.1, digits = c(2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S4:** Sub-microbial model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(two_pool_r_results_good[, c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S5:** Goodness of fit of Sub-microbial model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

@@

### Both models are calibrated against respiration rate, proteins and DNA concentration in the cells.

Both models are calibrated against all data we measured. In addition to model parameters reported in tables above, conversion factors between proteins/DNA concentration and microbial biomass are estimated. For Monod model, initial concentration of microbial biomass is defined as an initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and microbial biomass. For Sub-microbial model, initial concentration of Structures is defined similarly - i.e. initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and Structures. However, initial concetration of Reserves cannot be defined because initial concentration of proteins is not know. Therefore, one more parameter is defined. This parameter is denoted as RS_0 in the respective function and define initial concentration of Reserves per unit of Structures. This parameter has to be fixed across treatments to reflect the reality.
In addition, Monod model is calibrated against respiration rate and proteins or DNA concentration separately. When calibrated againnst protein concentration, initial concentration of microbial biomass is estimated and fixed across treatments.

#### Monod model

```{r Monod model - biomass components, eval=F, echo=T}
#Proteins
#Loading the function 
source("../Rscripts/monod_prot.R")
source("../Rscripts/monod_prot_fix.R")
#across all treatments
monod_prot1<-monod_prot(data=d, FACT = 4)
###############################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###############################################################
#for different structures
monod_prot2<-monod_prot(data=d, FACT = 2)
monod_prot2_fix<-monod_prot_fix(data=d, FACT = 2)
#for different substrates
monod_prot3<-monod_prot(data=d, FACT = 1)
monod_prot3_fix<-monod_prot_fix(data=d, FACT = 1)
#for each separately 
monod_prot4<-monod_prot(data=d, FACT = 3)
monod_prot4_fix<-monod_prot_fix(data=d, FACT = 3)
###############################################################
stopImplicitCluster()
###############################################################
monod_prot1$goodness
monod_prot2_fix$goodness
monod_prot3_fix$goodness
monod_prot4_fix$goodness

@@

kable(t.s4.1, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S6:** Monod model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_all_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S7:** Goodness of fit of Monod model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 3) %>%
  group_rows("Structure", 4, 6) %>%
  group_rows("Substrate", 7, 9) %>%
  group_rows("Each", 10, 12)

@@

kable(t.s4.2, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S8:** Monod model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_prot_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S9:** Goodness of fit of Monod model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8)

@@

kable(t.s4.3, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S10:** Monod model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_DNA_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S11:** Goodness of fit of Monod model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8)

@@

#### Sub-microbial model

Not all model parameters vary significantly across treatments. Parameters that doesn't show any variability across treatments are fixed.

```{r Sub-microbial model calibration, eval=F, echo=T}
#Loading the function 
source("./Rscripts/two_pool_allx.R")
source("../Rscripts/two_pool_allx_fix.R")
#across all treatments
two_pool_allx_out1<-two_pool_allx(data=d, FACT = 4)

@@

# write.csv(two_pool_all_results_par, file = c("./two_pool_all_results_par.csv"), row.names = F)
# write.csv(two_pool_all_results_good, file = c("./two_pool_all_results_good.csv"), row.names = F)

@@

kable(t.s6.1, digits = c(0, 2, 2, 2, 3, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ","v", 'k', 'f', 'm', 'Y', '$\\B_{R-Proteins}$', '$\\B_{S-Proteins}$', '$\\B_{S-DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S12:** Sub-microbial model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(two_pool_all_results_good[, c(1, 4:7)], digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S13:** Goodness of fit of Sub-microbial model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 3) %>%
  group_rows("Structure", 4, 6) %>%
  group_rows("Substrate", 7, 9) %>%
  group_rows("Each", 10, 12)
